---
title: "Data import"
output: github_document
---

Load libraries that you will need:

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in some data

Read in the litters dataset.

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

pups_df = read_csv("./data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```


## 'select'

Let's select some columns.

```{r}
select(litters_df, group, litter_number)
select(litters_df, group, gd0_weight, gd18_weight)

```

Colon keeps columns between two variables
Starts with, contains
Minus signs says every variable except that one
Can use equal sign to rename variables or use 'rename'
Last line relocates 'litter_number' to be the first variable, can also use relocate function

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
select(litters_df, group, starts_with("pups"))
select(litters_df, -litter_number)

select(litters_df, GROUP = group, litter_number)

rename(litters_df, GROUP = group)

select(litters_df, litter_number, everything())

relocate(litters_df, litter_number)
```

Using the Pups data set, select litter number, sex, PD ears
Can use either formula here.

```{r}
pups_df
select(pups_df, litter_number, sex, pd_ears)
select(pups_df, litter_number:pd_ears)
```


