---
title: "Data import"
output: github_document
---

Load libraries that you will need:

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```

## Read in some data

Read in the litters dataset.

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## Take a look at the data

Printing in the console.

```{r}
litters_df
```

Check first 6 rows and last 6 rows of the dataset
```{r}
head(litters_df)
tail(litters_df)
```

Skimr feature tells you how many missing values there are and some basic descriptive statistics. It also prints out a histogram of numeric variables.
```{r}
skimr::skim(litters_df)
```

In the console, type "view(litters_df)" but avoid putting it in a code chunk the R markdown document.

Note:
<dbl> is a specific type of number variable

## Options to read_csv

Check out '?read_csv' for more information, opens help file for the read_csv function

Skip the first 10 rows, don't use first row as column names

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
```

Define all missing values as NA

```{r}
litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", ".", 999))
```

Column types - use "col_types"

## Other file formats

Read in an excel file, using readxl package.

```{r}
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
```

Read in range of data cells

```{r}
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
mlb_df
```

Check out '?read_excel' for more information, opens help file for the read_csv function

Read in a SAS file, using haven package.

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df = janitor::clean_names(pulse_df)
pulse_df
```

## Comparison with Base R

what about 'read.csv' ...?

this is part of base R (not RStudio), and it is much harder to look at

```{r}
litters_base = read.csv("./data/FAS_litters.csv")
litters_readr = read_csv("./data/FAS_litters.csv")

litters_base
litters_readr
```

