---
title: "Tidy Data"
output: github_document
---

Load libraries that you will need:

```{r setup}
library(tidyverse)
```

## pivot_longer

Load the PULSE data

```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names()
```

Wide format to long format ...

```{r}
pulse_data_tidy =
  pulse_data %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  )

pulse_data_tidy
```

Re-write, combine, and extend (to add a mutate step)

```{r}
pulse_data =
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) %>% 
  relocate(id, visit) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
```


## 'pivot_wider'

Make up some data

```{r}
analysis_result = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
  )

analysis_result %>% 
  pivot_wider(
    names_from = "time",
    values_from = "mean"
  )
```

## Binding rows

Using the Lord of the Rings data.

First step: import each table.

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")
```

Bind all of the rows together

```{r}
lotr_tidy =
  bind_rows(fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>% 
  relocate(movie) %>%
  pivot_longer(
    female:male,
    names_to = "gender",
    values_to = "words"
  )
```

## Joining datasets

Import and clean the FAS datasets.

```{r}
pups_df = 
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))

litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  relocate(litter_number) %>% 
  separate(group, into = c("dose", "day_of_tx"), sep = 3)
```

Next up, time to join them!
Litter number will be the key.

```{r}
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number") %>% 
  arrange(litter_number) %>% 
  relocate(litter_number, dose, day_of_tx) 
```

## Learning assessments

In the litters data, the variables gd0_weight and gd18_weight give the weight of the mother mouse on gestational days 0 and 18. Write a data cleaning chain that retains only litter_number and these columns; produces new variables gd and weight; and makes gd a numeric variable taking values 0 and 18.

```{r}
litters_df =
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(litter_number, gd0_weight, gd18_weight) %>% 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd",
    values_to = "weight") %>% 
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))

```

First dataset has responses to a question about operating systems, and the second has responses to questions about degree program and git experience. Write a code chunk that imports and cleans both datasets, and then joins them.

```{r}
surv_os = read_csv("./data/surv_os.csv") %>% 
  janitor::clean_names() %>%
  rename(uni = what_is_your_uni) %>% 
  rename(os = what_operating_system_do_you_use)

surv_program = read_csv("./data/surv_program_git.csv") %>% 
  janitor::clean_names()%>%
  rename(uni = what_is_your_uni) %>% 
  rename(degree = what_is_your_degree_program) %>% 
  rename(git_experience = which_most_accurately_describes_your_experience_with_git)

surv_df = 
  left_join(surv_os, surv_program, by = "uni") %>% 
  arrange(uni) %>% 
  relocate(uni, os, degree) 

surv_df_2 =
  inner_join(surv_os, surv_program)
```

